---
title: "NRL DataJam 2019 - Penalty Probability"
output: html_notebook
---

## Penalty Probability 1st Pass

```{r}

library(dplyr)
library(ggplot2)
library(readr)
library(data.table)


```


```{r}

df_trx <- read_delim('C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\data/2019 Event Data/trxs.txt',delim='\t')

# df_trx %>% head()

```

#### Check out one game

```{r}

df_trx %>% filter(MatchId==18111011)

```

#### How many matches are there?

```{r}

cat('Total matches over 2018-2019:',length(unique(df_trx$MatchId)))

```

#### Subset only events where the penalty occurred and a decision was made  
- EventName contains 'Penalised', contains qualifier2name 'Taking penalty decision'  
- Qualifier2 contains the decision (ie. take penalty goal, kick for touch or tap)  
- Use OppositionID for penalty decision - since this event refers to the team that conceded the penalty  

#### High Level stats over all matches

```{r}

all_penalties = df_trx %>% 
  filter(EventName %like% 'Penalised')

print(nrow(all_penalties))

```

5900 penalty decisions made in entire last 2 seasons

#### How many penalties per match?

```{r}
# Penalties per match
all_penalties %>% group_by(MatchId) %>% 
  summarise(penalties_per_match = n())
```

#### What's the average number of penalty decisions per match (both teams)?

```{r}

# Average for all matches (both teams)
mean((all_penalties %>% group_by(MatchId) %>% 
  summarise(penalties_per_match = n()))$penalties_per_match)

```

#### What's the ratio of penalty decisions (kick for goal, kick for touch, tap)?
- not sure how the 2nd penalty and penalty try factor into the rules, its a small number so maybe I'll just exclude them    
- of course the highest proportion goes to kick for touch at 57% and 29% tap  
- 12% of penalties, the team opted for the penalty goal  
- ~9% of all kicks for goal missed the mark   
- 0.6% of kick for touch missed  

```{r}

data.frame(table(all_penalties$Qualifier2),prop.table(table(all_penalties$Qualifier2)))

```


```{r}

all_penalties_attacking <- all_penalties %>% filter(ZonePlayer <=28)

data.frame(table(all_penalties_attacking$Qualifier2),prop.table(table(all_penalties_attacking$Qualifier2)))


```
```{r}
all_penalties


```


```{r}

kick_forgoal_perc <- all_penalties %>% group_by(ZonePlayer,ChannelPlayer,SectionPlayer) %>% 
  summarise(kickforgoal_perc = sum(Qualifier2 %like% 'Kick for goal')/n())

kick_forgoal_perc <- kick_forgoal_perc %>% ungroup() %>% 
  mutate(SectionPlayer=as.factor(SectionPlayer),
         ChannelPlayer=as.factor(ChannelPlayer))

ord_channels <- c("SR","NR","R","C","L","NL","SL")
ord_sections <- c("-0", "-10", "-20","-30","-40","-50","-60","-70","-80","-90", "-100","-110")

kick_forgoal_perc$SectionPlayer <- factor(kick_forgoal_perc$SectionPlayer, levels = ord_sections,ordered = T)
kick_forgoal_perc$ChannelPlayer <- factor(kick_forgoal_perc$ChannelPlayer, levels = ord_channels,ordered = T)

ggplot(kick_forgoal_perc,aes(x= SectionPlayer,y=ChannelPlayer)) + 
  geom_tile(aes(fill = kickforgoal_perc)) +scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  geom_text(aes(label = round(kickforgoal_perc, 2)))


```



#### Let's look all penalty goals and conversions and see what the accuracy is for all teams for different areas of the field
- EventName contains 'Penalty goal' (OK or missed)
- EventName contains 'Conversion' (OK or missed)
- Use XmPlayer and YmPlayer (normalised to player tagged to the event)

```{r}

all_pen_conv <- df_trx %>% 
  filter(EventName %like% 'Penalty goal -' | EventName %like% 'Conversion -')

data.frame(table(all_pen_conv$EventName),prop.table(table(all_pen_conv$EventName)))

```

As seen above, most of the examples we have are of conversions. But we'll just make the assumption that the difficulty/accuracy should be similar for both penalty goals and conversions. Note: What is 'Conversion - 8pt OK' - found on google, potential 8pt try if try scorer is fouled after scoring try.  

#### Add the conversions and penalties together for overall accuracy

```{r}

cat('Overall accuracy of goal kicks: ', (1960+2+691)/nrow(all_pen_conv))

```

#### How many unique players have taken kicks in the last 2 seasons?

```{r}

length(unique(all_pen_conv$PlayerId))

```

#### Average number of kicks made by each kicker?
- Very high variability in the number of kicks. there is definitely 1 main kicker and couple of backup kickers in each team.
- Therefore, let's just simplify things and do Penalty Probability at the Team level instead of the kicker level.

```{r}
all_pen_conv %>% group_by(ClubId,PlayerId) %>% 
  summarise(total_kicks = n())
```


```{r}

ggplot((all_pen_conv %>% group_by(ClubId,PlayerId) %>% 
  summarise(total_kicks = n())),aes(y=total_kicks)) +geom_boxplot()

```


#### Average number of kickers per team in the last 2 seasons?

```{r}

mean((all_pen_conv %>% group_by(ClubId) %>% 
  summarise(num_kickers = n_distinct(PlayerId)))$num_kickers)

```



#### Now get accuracy for each zone on field  
- group by ZonePlayer

* Need to correct in the Section numbers: Zones 78-84 seem to have been mapped to section '-0', but they should be section '-110'

```{r}

acc_per_zone <-  all_pen_conv %>% 
  group_by(ZonePlayer,ChannelPlayer,SectionPlayer) %>%
  summarise(total_kicks = n(),
            zone_accuracy = sum(EventName %like% 'OK')/n())

# Correct an error in section numbers
acc_per_zone <- acc_per_zone %>% 
  filter(!ZonePlayer %in% seq(78,84))

acc_per_zone <- acc_per_zone %>% ungroup() %>% 
  mutate(SectionPlayer=as.factor(SectionPlayer),
         ChannelPlayer=as.factor(ChannelPlayer))

acc_per_zone

```

Looks like most zones that conversions or penalty goals are kicked from from are from ~40m up to ~10m out from try line. There seem to be 4 kicks that are in zones 79-84, which doesn't make sense as these are inside the in-goal area (maybe mistakes in the data).  


#### Scatterplot of where kicks are made from
- Conversions have a nice curve in 10-30m in front of goal posts  
- Some dirtiness in the data, where the xy-coordinates were not normalised properly. Also conversion kicks that are inside ingoal area

```{r}

ggplot(all_pen_conv,aes(x= XmPlayer,y=YmPlayer,color=EventName)) + geom_point()

```

#### Heatmap of accuracies by zone  
- Probs ignore anything smaller than section -50

```{r}
ord_channels <- c("SR","NR","R","C","L","NL","SL")
ord_sections <- c("-0", "-10", "-20","-30","-40","-50","-60","-70","-80","-90", "-100","-110")

acc_per_zone$SectionPlayer <- factor(acc_per_zone$SectionPlayer, levels = ord_sections,ordered = T)
acc_per_zone$ChannelPlayer <- factor(acc_per_zone$ChannelPlayer, levels = ord_channels,ordered = T)

ggplot(acc_per_zone %>% filter(total_kicks>=5),aes(x= SectionPlayer,y=ChannelPlayer)) + 
  geom_tile(aes(fill = zone_accuracy)) + scale_fill_gradient(low = "#CFF4D2", high = "#205072") + #scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  geom_text(aes(label = round(zone_accuracy, 2)))

```

#### Total Kick Attempts by Zone

```{r}

ggplot(acc_per_zone,aes(x= SectionPlayer,y=ChannelPlayer)) + 
  geom_tile(aes(fill = total_kicks)) +scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  geom_text(aes(label = total_kicks))

```

#### Accuracy per Club per Zone for Tableau
- Grouping by club ensures we also take into account the kicker accuracy 

```{r}

team_acc_per_zone <-  all_pen_conv %>% 
  group_by(ZonePlayer,ChannelPlayer,SectionPlayer,ClubId) %>%
  summarise(total_kicks = n(),
            zone_accuracy = sum(EventName %like% 'OK')/n())

# Remove zones that are over the try line
team_acc_per_zone <- team_acc_per_zone %>% 
  filter(!ZonePlayer %in% seq(78,84))

team_acc_per_zone <- team_acc_per_zone %>% ungroup() %>% 
  mutate(SectionPlayer=as.factor(SectionPlayer),
         ChannelPlayer=as.factor(ChannelPlayer))

team_acc_per_zone

```

Which teams/zones have fewer than 5 total kicks in the whole dataset?
- as expected low numbers 

```{r}
team_acc_per_zone %>% filter(total_kicks<5)
```

#### Or Should We Do Club Accuracy Multiplied by Overall Zone Accuracy?
- However, this artificially reduces the actual accuracy measure for every team
 
- Best Goal Kicking Club: Penrith Panthers 
- Worst Goal Kicking Club: West Tigers

```{r}

acc_per_club <-  all_pen_conv %>% 
  group_by(ClubId) %>%
  summarise(total_kicks = n(),
            club_accuracy = sum(EventName %like% 'OK')/n()) %>% 
  arrange(club_accuracy)

acc_per_club

```

#### Add XY coords to table

```{r}

team_acc_per_zone <- team_acc_per_zone %>% 
  mutate(ZoneX = case_when(ChannelPlayer=="SL" ~ 10,
                           ChannelPlayer=="NL" ~ 20,
                           ChannelPlayer=="L" ~ 30,
                           ChannelPlayer=="C" ~ 40,
                           ChannelPlayer=="R" ~ 50,
                           ChannelPlayer=="NR"~ 60,
                           ChannelPlayer=="SR"~70),
         ZoneY = -as.integer(as.character(SectionPlayer)))


```
 
 
#### Save datasets/outputs

```{r}

# all_pen_conv %>% write_csv("C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\output\\penalties_conversions.csv")
# 
# acc_per_zone %>% write_csv("C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\output\\kick_acc_per_zone.csv")

team_acc_per_zone %>% write_csv("C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\output\\club_kicking_acc_per_zone.csv")

```



