---
title: "NRL DataJam 2019 - Try Probability"
output: html_notebook
---

## Try Probability 1st Pass

```{r}

library(dplyr)
library(ggplot2)
library(readr)
library(data.table)


```


```{r}

df_trx <- read_delim('C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\data/2019 Event Data/trxs.txt',delim='\t')

df_trx %>% head()

```

#### Have a look at where the sets start in 1 game
- In the data, at the start of the half, the kickoff and subsequent kick chase seems to be counted as a set, so make sure we remove that from potential try-scoring sets  
- The sets seem to start where event name is 'Start set complete ...' or 'Start set incomplete ...'  


```{r}

df_trx %>% filter(MatchId==18111011) %>% filter(EventName %like% 'Start set complete' | EventName %like% 'Start set incomplete') 

```

#### How many matches are there?

```{r}

cat('Total matches over 2018-2019:',length(unique(df_trx$MatchId)))

```

#### How many total sets in each match?

```{r}

all_sets <- df_trx %>% 
  filter(EventName %like% 'Start set complete' | EventName %like% 'Start set incomplete') 

print(nrow(all_sets))

```

30,008 sets in last 2 seasons

#### How many sets per match?

```{r}
# Sets per match
all_sets %>% group_by(MatchId) %>% 
  summarise(sets_per_match = n())
```


#### What's the average number of sets per match (both teams combined)?

```{r}

# Average for all matches (both teams)
mean((all_sets %>% group_by(MatchId) %>% 
  summarise(sets_per_match = n()))$sets_per_match)

```

#### Get every set of 6 in every match
- Every set of 6
- Extract X-Y coordinates at start of set - use XmPossession, YmPossession (normalised to team with posession during set)
- Extract result of set (Try or Not) - every important detail about the set is encoded in the row where eventname = 'start set complete/incomplete'
  - Qualifier3 gives the set outcome (try, turnover, kick, penalty etc)
  - use Qualifier1 to get possession source (eg. if it started from a penalty)

#### What's the result of a set of 6?

```{r}

data.frame(table(all_sets$Qualifier3),prop.table(table(all_sets$Qualifier3)))

```

- 9% of all sets result in a Try  
- 65% of sets end in a kick, 18% of sets end in a tackle turnover  



#### Let's look at probability of Try all teams for different areas of the field
- XmPossession, YmPossession (normalised to team with posession during set)
- group by ZonePossession, ChannelPosession and SectionPosession

```{r}

prob_per_zone = all_sets %>% 
  group_by(ZonePossession,ChannelPossession,SectionPossession) %>%
  summarise(num_sets = n(),
            zone_prob = sum(Qualifier3 =='Try')/n())

# Correct an error in section numbers
prob_per_zone <- prob_per_zone %>% 
  mutate(SectionPossession = ifelse(ZonePossession %in% seq(78,84), '-110',SectionPossession))

prob_per_zone <- prob_per_zone %>% ungroup() %>% 
  mutate(SectionPossession=as.factor(SectionPossession),
         ChannelPossession=as.factor(ChannelPossession))

prob_per_zone

```

#### Scatterplot of where sets start
- plot doesn't tell us a whole lot

```{r}

ggplot(all_sets,aes(x= XmPossession,y=YmPossession,color=Qualifier3)) + geom_point(alpha=0.3)

```

#### Heatmap of try probability by zone  

```{r}
ord_channels <- c("SL","NL","L","C","R","NR","SR")
ord_sections <- c("IG-", "-10", "-20","-30","-40","-50","-60","-70","-80","-90", "-100","-110")

prob_per_zone$SectionPossession <- factor(prob_per_zone$SectionPossession, levels = ord_sections,ordered = T)
prob_per_zone$ChannelPossession <- factor(prob_per_zone$ChannelPossession, levels = ord_channels,ordered = T)

ggplot(prob_per_zone,aes(x= SectionPossession,y=ChannelPossession)) + 
  geom_tile(aes(fill = zone_prob)) +scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  geom_text(aes(label = round(zone_prob, 2)))

```

- Note: Need to double check that the zones and sections have been normalised properly

#### Total Number of Sets by Zone

```{r}

ggplot(prob_per_zone,aes(x= SectionPossession,y=ChannelPossession)) + 
  geom_tile(aes(fill = num_sets)) +scale_fill_continuous(high = "#132B43", low = "#56B1F7")+
  geom_text(aes(label = num_sets))

```
 
#### Save datasets/outputs

```{r}

all_sets %>% write_csv("C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\output\\sets_outcomes.csv")

prob_per_zone %>% write_csv("C:\\Users\\martin.cheung\\OneDrive - Accenture\\NRL DataJam 2019\\output\\try_prob_per_zone.csv")

```



